Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-12-18 paper.tex
updated manuscript; added ndmg bits

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, Pierre Bellec,\\
        Remi Bernard, Sorina Camarasu-Pop, Fr\'ed\'eric Cervenansky, Samir Das, \\
        Rafael Ferreira da Silva, Guillaume Flandin, John Flavin, Pascal Girard, \\
         Krzysztof J. Gorgolewski, Charles G. Guttmann, Nathaniel Kofalt, Pierre-Olivier Quirion,\\
         Pierre Rioux, Marc-\'Etienne Rousseau, Gunnar Schaeffer, Alan C. Evans}

@@

The core component of the manifest is a flexible command-line
description represented as a simple string template complying to the
syntax of the \texttt{bash} UNIX shell. \texttt{bash} is the default
shell on the major Linux distributions and OS X. The command-line
template may contain placeholders for input and output values, called
value keys. The command line may actually
encompass several commands separated, e.g., by semicolumns, pipes
(\texttt{|}) or ampersands (\texttt{\&}). Such a flexible command-line
specification is meant to facilitate the embedding of minor mundane
operations on the command line, for instance directory creation, input
decompression or output archival.

@@

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} or \texttt{File}. \texttt{File} may
also represents a directory. Types can be restricted to a specific set
of values (\texttt{String} and \texttt{Number} only), to a specific
range (\texttt{Number} only), to integers or to booleans (\texttt{Number}
only).

@@

The Boutiques validator checks conformance of JSON manifests to the
Boutiques schema using a basic JSON validator. It also performs the
following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, value keys are not contained within each other (which would
puzzle substitution), output path templates are unique (to avoid
results to overwrite each other), inputs of type Flag have a
command-line flag, are optional and are not lists, number and list
constraints are sensible (e.g. min is lower than max), the default
value of enum types is part of the enum, an input cannot both require
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, and
one-is-required groups should never have required members. The
Boutiques validator is meant for application developers; platforms
should rely on regular JSON validation to avoid depending on a
third-party tool.

@@

@@

    \hline
    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    FSL tools &&&&&&\\
    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
    \end{tabular}
  \vspace*{0.5cm}
  \begin{tabular}{l|ccc|ccc}
    \hline
    \rowcolor[gray]{0.9}
    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
    \rowcolor[gray]{0.9}
                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
    \hline
    Niak pre-processing             &\cellcolor{gray!75} &            &         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
    FSL tools &&&&&\\
    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &             &&\\
    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &             &&\\
  \end{tabular} 
  \caption{Applications available in Boutiques with manifest features used.}
  \label{table:applications}
\end{table}

\section{Discussion}

@@

CWL has a formal command-line definition while \boutiques is more
flexible. CWL specifies command lines using an array of executable and
arguments whereas \boutiques only uses a string template. \boutiques'
template approach may create issues in some cases, but it also allows
developers to add simple operations to an application without having
to write a specific wrapper. For instance, a \boutiques command line
may easily include input decompression using, e.g., the tar command in
addition to the main tool command. On the contrary, CWL applications
may only have a single executable. Importantly, \boutiques' template
system allows supporting configuration files (see
Section~\ref{sec:configuration-files}).

>>>>

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, Pierre Bellec,\\
        Remi Bernard, Sorina Camarasu-Pop, Fr\'ed\'eric Cervenansky, Samir Das, \\
        Rafael Ferreira da Silva, Guillaume Flandin, John Flavin, Pascal Girard, \\
        Krzysztof J. Gorgolewski, Gregory Kiar, Charles G. Guttmann, Nathaniel Kofalt, \\
        Pierre-Olivier Quirion, Pierre Rioux, Marc-\'Etienne Rousseau, Gunnar Schaeffer, Alan C. Evans}

@@

The core component of the manifest is a flexible command-line
description represented as a simple string template complying to the
syntax of the \texttt{bash} UNIX shell. \texttt{bash} is the default
shell on the major Linux distributions and OS X. The command-line
template may contain placeholders for input and output values, called
value keys. The command line may actually
encompass several commands separated, e.g., by semicolons, pipes
(\texttt{|}) or ampersands (\texttt{\&}). Such a flexible command-line
specification is meant to facilitate the embedding of minor mundane
operations on the command line, for instance directory creation, input
decompression or output archival.

@@

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} %, \texttt{Enum},
or \texttt{File}. \texttt{File} may
also represent a directory. Types can be restricted to a specific set
of values (\texttt{String} %, \texttt{Enum},
and \texttt{Number} only), to a specific
range (\texttt{Number} only), to integers or to booleans (\texttt{Number}
only) %, or to an arbitrary but specified set of values (\texttt{Enum} only)
.

@@

The Boutiques validator checks conformance of JSON manifests to the
Boutiques schema using a basic JSON validator. It also performs the
following checks that cannot be easily implemented in JSON schema:
value keys are unique among inputs, input and output identifiers are
unique, input and output value keys are all included in the command
line, value keys are not contained within each other (which would
puzzle substitution), output path templates are unique (to avoid
results to overwrite each other), inputs of type Flag have a
command-line flag, are optional and are not lists, number and list
constraints are sensible (e.g. min is lower than max), the default
value of enum types is part of the enum, an input cannot both require
\note{greg}{should Enum be capitalized? It is earlier in the document}
and disable another input, required inputs cannot require or disable
other parameters, group member identifiers must correspond to existing
inputs and cannot appear in different groups, mutually exclusive
groups cannot have members requiring other members, and
one-is-required groups should never have required members. The
Boutiques validator is meant for application developers; platforms
should rely on regular JSON validation to avoid depending on a
third-party tool.

@@

\paragraph{ndmg.} The NeuroData MRI to Graphs one-click connectome estimation
pipeline~\cite{kiar2016ndmg}, developed in Python and leveraging FSL, was
exported to \boutiques and is available at \url{https://github.com/neurodata/boutiques-tools}.

@@

    \hline
    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    FSL tools &&&&&&\\
    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
    ndmg                &                   &       \cellcolor{gray!75}            &                   &          \cellcolor{gray!75}         &      \cellcolor{gray!75}             &                   &                   \\
    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
    \end{tabular}
  \vspace*{0.5cm}
  \begin{tabular}{l|ccc|ccc}
    \hline
    \rowcolor[gray]{0.9}
    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
    \rowcolor[gray]{0.9}
                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
    \hline
    Niak pre-processing             &\cellcolor{gray!75} &            &         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
    FSL tools &&&&&\\
    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
    FSL Nipype interfaces           &                    &            &         &              &&\\
    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &              &&\\
    ndmg                            &\cellcolor{gray!75} &            &         &              &&\\
    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &              &&\\
  \end{tabular} 
  \caption{Applications available in Boutiques with manifest features used.}
  \label{table:applications}
\end{table}
\note{greg}{@tristan g: chris helped me verify that ndmg also works with singularity, but we have not shared
the image anywhere. Would it be worth me publishing that image for this table?}
\section{Discussion}

@@

CWL has a formal command-line definition while \boutiques is more
flexible. CWL specifies command lines using an array of executables and
arguments whereas \boutiques only uses a string template. \boutiques'
template approach may create issues in some cases, but it also allows
developers to add simple operations to an application without having
to write a specific wrapper. For instance, a \boutiques command line
may easily include input decompression using, e.g., the tar command in
addition to the main tool command. On the contrary, CWL applications
may only have a single executable. Importantly, \boutiques' template
system allows supporting configuration files (see
Section~\ref{sec:configuration-files}).
